Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-08-25 documents/ovando-etal-assessing-global-fisheries.Rmd
passing back to squid

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results and therefore, high uncertainties in our understanding of global fisheries. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a new flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. We show that our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(viridis)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(here)
extrafont::loadfonts()

@@

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fishery, 6.2% are underfished ,and 34.2% are overfished [@fao2020]. While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment. Although revised periodically over time, the method has kept revisions to a minimally necessary levels for the purposes of temporal consistency and comparability. With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert elicitation. While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].

Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance such as catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

@@

However, @free2020 demonstrated that these catch-only models can often produce both imprecise and biased estimates of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but other data sources such as surveys and expert opinion suggest this region should have more depleted fisheries than Northeast Pacific. The Costello et al. 20162 model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).

@@

We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@melnychuk2017], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model. `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.

@@

```{}

@@

@@

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world. Numerous efforts have sought to resolve this problem through development of new modeling methods. These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment. However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes. While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.

What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends of course on the needs of those using the assessment outputs. It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient. We have stated that addition of the globally available datasets compiled here did not substantially improve estimates of stock status, but in some instances using these data did provide some improvement over use of catch-only style methods. However, none of the data sources used here came close to the benchmark performance provided by our fits to the RAM data, either in overall performance or in capturing the geographic distribution of stock status. While it is unreasonable to expect any global-scale dataset to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space. That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.

@@

Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%. In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot). Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance. This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.

@@

Another avenue would be to prioritize the development of a global repository for length and age composition data. The data-limited assessment world has recently been developing along two almost parallel tracks. Catch-based methods have most commonly been employed in efforts to understand global fisheries, and on occasion for management of stocks mandated to manage catch but lacking sufficient data for a formal stock assessment [@berkson2015]. Along with general performance questions around catch-only models, no matter what they require a complete, or at least near complete timeseries of captures. This rules catch-based methods out for many small scale fisheries around the world. As a result, a growing number of methods have focused on using length-composition data to provide estimates of stock status, including measures of fishing mortality rates [@hordyk2016; @prince2019; @rudd2017]. These methods have a distinct advantage for many fisheries in that they can in theory function with as little as one year of data. As a result, many fisheries around the world are enacting programs to collect and utilize length composition data for fisheries assessment.

@@

All analysis were conducted in the R programming language [@rcoreteam2019]. Unless otherwise stated all model fitting was conducted using stan [@carpenter2017], implemented through template model builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The `sraplus` package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).

## `sraplus`

@@

```{r}
cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))

knitr::kable(
  cd,
  "latex",
  align = "l",
  booktabs = TRUE,
  escape = F,
  row.names = F,
  caption = "Estimated parameters, abbreviations, and default priors"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

@@

## Priors and Data

We briefly describe the construction of priors and nature of available data. All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation]. These models are intended as a reasonably robust means of translating available data into usable prior for our model. Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlike that, given the same data, and alternative model would be unlikely to perform much better than those presented here XX this might be too much defenseXX

## Swept Area Ratio Priors

Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018]. While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and a heuristic approach.

For the heuristic approach, we borrow from the original methods of @martell2013 as presented in @andersno2017, and define a heuristic on initial and terminal depletion where

```r
        if (first(catch) / max(catch) < 0.2){

        initial_depletion = 0.7

        } else {

        initial_depletion = 0.4

        }

        if (last(catch) / max(catch) > 0.5){

        terminal_depletion <- 0.6

        } else {

        terminal_depletion <- 0.2

        }

# Supplemental Figures

@@

>>>>

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methdological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that catch data by themselves cannot provide reliable estimates of fishery health. Our ability to obtain reliable estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of key data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(viridis)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(here)
library(sraplus)
extrafont::loadfonts()

@@

# rishi notes

* more assertive on what the data needs really are
  * where to apply globally 
* For places with less coverage, what would we actually recommend there 
* what do we know about catchabiliy
- e.g. 51, fleets became industrial 
  - hammer down on that's the issue with the effort 
  - how do we translate effort into f bey area
  - here's

@@

The United Nations Sustainable Development Goal 14, related to "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". A substantial part of achieving this goal rests on ensuring that marine fisheries are managed to levels that ensure the conservation and sustainable use of ocean ecosystems. Achieving this goal depends in part on our ability to effectively track the status of global marine fish stocks. The Food and Agriculture Organization of the United Nations' (FAO) State of World Fisheries and Aquaculture (SOFIA) report is the standard source for tracking this important metric. As of the most recent report, the FAO estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished ,and 34.2% are overfished [@fao2020]. While foundational, the SOFIA assessment, designed in the 1970s based on the then available data and methods, adopts a hybrid approach that combines results of numerical models, indicator-based trends and educated expert judgment. Although revised periodically over time, the method has kept revisions to a minimally necessary levels for the purposes of temporal consistency and comparability. With the surge in data availability and models designed for data-limited stocks, new global assessment methods are needed to meet the new demand for estimation and tracking progress towards the SDG goals.

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts often are well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018].

Formal stock assessment require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist, and would be prohibitively expense to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

@@

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status (B/B~MSY~). These issues become apparent when we consider some of the macro-level predictions made by these models. The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020]. A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.

xx spend time on this tomorrow xx
Examining stock assessment data from RAM from the NE Pacific, the NE Atlantic and the Mediterranean, the NE Pacific is in the best state (mean RAM b/b~MSY~ = 1.8), NE Atlantic (mean RAM b/b~MSY~ =1.3) and the Mediterranean the worst (mean RAM b/b~MSY~ =0.5).  We do not have formal stock assessment data from S. and SE Asia (Western Central Pacific and Eastern Indian Ocean), but other data sources such as surveys and expert opinion suggest this region should have more depleted fisheries than Northeast Pacific. The Costello et al. 2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and SE Asia seem to be overoptimistic, and the NE Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in SE Asia estimated as doing better than the NE Atlantic or NE Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)).

@@

xx move some of this to intro and chop downXX
We use a flexible surplus-production based stock assessment package, [`sraplus`](www.github.com/danovando/sraplus) to demonstrate how different sources data can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. The FAO's catch statistics remain the most widely accessible and used source of data for the assessment of global fishery sustainability. However, catches are not the only broadly available source of fishery data. We augment the FAO's catch statistics with broadly distributed data on fishing effort [@rousseau2019], strength of management [@melnychuk2017], trawl-fishing footprints [@amoroso2018], and life history [@thorson2017c]. `sraplus` provides a flexible framework for integrating these data into valid statistical model. `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.

@@

XX add stock assessment estimates for both regions? XX Maybe just do matching stocks?

@@

 xx dig into this, which stocks are in that region? XX

```{r}
a  = ram_comp_data %>% 
  filter(fao_area_code == 61)

unique(a$common_name)
```

@@

<!-- Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world. Numerous efforts have sought to resolve this problem through development of new modeling methods. These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment. However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes. While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.   -->

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world. While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone, at the global level no combination of readily available data was able to substantially and reliably clarify our picture of global fisheries status.

What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends of course on the needs of those using the assessment outputs. It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient. We have stated that addition of the globally available datasets compiled here did not substantially improve estimates of stock status, but in some instances using these data did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe. However, none of the data sources used here came close to the benchmark performance provided by our fits to the RAM data, either in overall performance or in capturing the geographic distribution of stock status. While it is unreasonable to expect any global-scale data to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space. That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.

@@

Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 20%. In terms of data however, information on fishing mortality rates (F/F~MSY~) in the most recent year was almost as useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot). Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance (xx uep, since in some cases the index was flatish but the stock status was low, so a model fit just to the trend in the abundance gets the trend right but the absolute value wrong. having F/Fmsy data helps anchor that). This is because we varied both the length and quality of the index of abundance, and on average an even noisy estimate of recent F/F~MSY~ may be more valuable than a short and noisy abundance index.

@@

Another avenue would be to prioritize the development of a global repository for length and age composition data. The data-limited assessment world has recently been developing along two almost parallel tracks. Catch-based methods have most commonly been employed in efforts to understand global fisheries, and on occasion for management of stocks mandated to manage catch but lacking sufficient data for a formal stock assessment [@berkson2015]. Along with general performance questions around catch-only models, no matter what they require a complete, or at least near complete timeseries of captures. This rules catch-based methods out for many small scale fisheries around the world. As a result, a growing number of methods have focused on using length-composition data to provide estimates of stock status, including measures of fishing mortality rates [@hordyk2016; @prince2019; @rudd2017]. These methods have an advantage for many fisheries in that they can in theory function with as little as one year of data. As a result, many fisheries around the world are enacting programs to collect and utilize length composition data for fisheries assessment. While length-based assessments come with a host of assumptions and pitfalls, properly implemented in some fisheries this may provide an overlooked source of fisheries data at a global scale.

@@

All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [eddelbuettel2011] and stan [@carpenter2017], implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The `sraplus` package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries).

## Model Overview

The core of `sraplus` is a Pella-Tomlinson [@pella1969] production model  in the manner of @winker2018. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the potential data evaluated here will support. The purpose of `sraplus` is not to make substantial improvements in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status. Detailed examples of how to operate the code of the model can be found [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus). Here we describe the overall methods used to produce the results in this paper and highlight key details.

`sraplus` can be used in two general ways: as a mechanism for revealing implicit priors on fishery status through a sample-importance-resampling algorithm, and as a means for fitting indices of abundance informed by priors. To illustrate we will walk through a series of ways that `sraplus` could be applied to a case study fishery.

Suppose that the only thing we observe from a fishery is a catch history. Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past (i.e. that catches have never been greater than biomass). We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history.  Beyond that though, baring additional information we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the current population must be greater than 0 and less than or equal to carrying capacity *k*.

What hapens when we impose informative priors on the population parameters? For example at a bare minimum, we know that carrying capacity must be less than infinity, and we can certainly make inferences based on knowledge from other better studied fisheries of similar species on what a potential upper bound on carrying capacity might be. Once we have generated informative priors on our population parameters, we have implicitly provided priors on the biomass and therefore status of the fishery over its entire history: if our priors on carrying capacity only extend a small fraction beyond the minimum value needed to ensure the existence of the population, our implicit prior is that the population is severely depleted. If our prior is that carrying capacity is orders of magnitude greater than the observed catches, then our implicit prior is that the population is relatively unexploited.

For our case study fishery, we can use `sraplus` as a filter to say, conditional on our priors, the catch history, and the population model, what is our implied prior on stock status of the fishery over time (see Section XX and Table.\@ref(tab:prior-tab) for descriptions of model parameters). While user defined priors on population parameters can be supplied, `sraplus` has a number of built in mechanisms for setting parameters based on the taxonomy of the species in question and the magnitude of the catch history. This form of analysis is often extended though to include an explicit prior on stock status in the most recent year [@walters2006].

For our case study, expert opinion tells us that the stock is likely between XX and XX% of carrying capacity in the most recent year. While we can use `sraplus` to include this addition information, doing so introduces a nontrivial statistical problem: we have now placed two priors on the same outcome (current stock status), one explicit and one implicit in our growth parameter priors.  The net effect of this is that once these priors are run through `sraplus`, our post-model-pre-data "estimate" of stock status will appear to have "updated", creating an illusion that we have learned something about stock status. Remember though that we have absolutely no data on which to update our prior on stock status beyond the minimum floor needed to ensure that the population exists.

`sraplus` implements an approximate solution to this problem. Rather than using combinations of implied and explicit priors, `sraplus` asks, given an explicit prior on recent stock status, what prior distributions of growth parameters produce our explicit prior on recent stock status, given the catch history and model assumptions. This helps resolve some of the statistical issues related to placing multiple priors on the same parameter, and makes clear where any information on recent stock status is coming from. To return to our case study then, uncorrected, combining our explicit priors on recent stock status with our population parameter priors reveals an implicit prior on recent stock status that spans XX to XX. Using our correction, the implicit prior on recent stock status more or less matches our explicit prior (Fig.XX).

While `sraplus` lets users supply their best estimate of recent stock status for a given fishery, the model also allows users to supply data on the swept area ratio (SAR , @amoroso2018) and/or fisheries management index (FMI, @melnychuk2017) scores for a given fishery. SAR is a measure of the trawling intensity in a given region, and as such for species susceptible to trawl gears can be an effective index of fishing pressure. FMI reflects expert opinion on the research, management, enforcement, and socioeconomic capacity of a given fishery, which together have significant associations with stock status. `sraplus` uses relationships between these SAR and FMI scores and observed F/F~MSY~  and B/B~MSY~ values to generate a posterior predictive distribution of F/F~MSY~ or B/B~MSY~ for the fishery in question, which serves as the priors for `sraplus`.

To continue our case study example, suppose that we no longer believe we have good data on current stock status, but do have a reasonable estimate of swept area ratio. `sraplus` now uses the posterior predictive F/F~MSY~ generated by the SAR value as the prior for recent F/F~MSY~ for the fishery in question. This process allows users to translate something that they can measure, SAR, into something that they perhaps cannot (F/F~MSY~). The model then determines the growth parameter priors that conditional on the catch history and model structure produce the SAR generated F/F~MSY~ prior

Up to this point we still have not fit the model to any data; sraplus is simply helping us visualize the implications of our priors in terms of stock status over time and relevant reference points such as MSY. In order to actually learn something new about the stock in question, we need to update our priors with data. `sraplus` allows for two methods for accomplishing this: the model can be fit to an observed index of abundance (for example a fishery independent survey), or fit to a catch-per-unit effort index calculated from the observed catch and an index of fishing effort. In the latter case, the model must make an assumption about the rate of efficiency increase of a unit of fishing effort over time. When data are available the model can be provided priors on recent fishery status derived from expert opinion or FMI/SAR values as well.

pull out example from SAR fits
```{r}
sar <- sraplus::sar %>% 
  filter(stockid %in% unique(ram_data$stockid))

ex <- ram_data %>% 
  filter(stockid == "PLAICIS")

ex %>% 
  ggplot(aes(year, b_v_bmsy)) + 
  geom_point()

# fit using catch heuristic

# fit using SAR

# fit using effort index

## Population Model

@@

```{r prior-tab}
cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))

knitr::kable(
  cd,
  "latex",
  align = "l",
  booktabs = TRUE,
  escape = F,
  row.names = F,
  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

## FMI and SAR Based Priors

All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [xx citation]. These models are intended as a reasonably robust means of translating available data into usable prior for our model. Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlike that, given the same data, and alternative model would be unlikely to perform much better than those presented here XX this might be too much defenseXX

### Swept Area Ratio Priors

Many of the current methods for estimating global stock status of unassessed stocks are based on inferring stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018]. While these catch-only methods have been shown to have series shortcomings [@free2020], as a point of reference we use two forms of catch-only priors: a statistical and CMSY, as presented implemented by @palomares2020.

For the CMSY based methods, we used a ported version of the CMSY model available [here](https://github.com/DanOvando/portedcmsy). For each stock we used all the default options provided by CMSY, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [boettiger2012]. Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.

# Supplemental Information

## Population Model

```{r si-prior-tab}
cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
  rbind(c("Shape parameter", "m", "$logn(1.01, 0.25)$")) %>%
  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
  rbind(c("Process Error", "$\\sigma_{proc}$", "$logn(.05,0.5)$")) %>% 
  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$"))

knitr::kable(
  cd,
  "latex",
  align = "l",
  booktabs = TRUE,
  escape = F,
  row.names = F,
  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

@@

@@

## SAR

## FMI

## Catch

fit_data <- fit$data %>% 
  group_by(stockid) %>% 
  filter(year == max(year))

pp <- posterior_predict(fit, draws = 1000, newdata = fit_data)
